---
title: "MVMR make plots"
author: "Marina Vabistsevits"
date: "01/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(readr)
library(dplyr)
library(data.table)
library(TwoSampleMR)
library(ggplot2)
library(cowplot)
library(wesanderson)
```

```{r message=F}
# set path for pre-calculated data, outside the code repo
# `local` / `remote` (reading data from RDSF)
currently_working_env = "local"
source("set_paths.R")
set_paths(currently_working_env)

# metafile
data_lookup<-read_csv(paste0("metadata/data_lookup.csv")) 
source("functions.R")

# breast cancer dataset
# 1126 full, 1127 ER+, 1128 ER-
breast_cancer_id <- "ieu-a-1126" 
bc_data <- stringr::str_split(breast_cancer_id, "-")[[1]][3]
```

```{r}
# specify category and palette to use

current_trait_category <- "glycemic_traits"
pal<-c(unname(yarrr::piratepal("pony")))

current_trait_category <- "reproductive_traits"
pal<-rev(unname(yarrr::piratepal("info2")))

current_trait_category <- "hormones"
pal <- wes_palette("Darjeeling1")[c(2,3,4,5)]

current_trait_category <- "hormones_splitsample"
pal <- wes_palette("Darjeeling1")[c(2,4,5)]

current_trait_category <- "physical_traits"
pal<-rev(unname(yarrr::piratepal("pony")))
```

## MVMR analysis 1: childhood BMI + adult BMI -> mediator

```{r message=F}
dat <- read_tsv(paste0(results_path, current_trait_category, 
                       "/merged/merged_mvmr_BMIs-to-", current_trait_category,".tsv"))

if (current_trait_category == "hormones"){ dat<-filter(dat, !grepl(",", outcome)) } # don't include testosterone subcats

p<-ggplot(dat, aes(y=exposure, x=b, label=outcome.full, shape=exposure, colour=trait_type)) +
  geom_errorbarh(aes(xmin=lo_ci, xmax=up_ci), height=.3) +
  geom_point(size=2)+
  scale_color_manual(values=pal)+
  scale_shape_manual(values = c(19,20)) +
  geom_vline(xintercept=0, linetype='longdash') +
  theme_minimal_hgrid(10, rel_small = 1) +
  facet_wrap(~outcome.full, ncol=1)+
  labs(color = "",y = "", x = "Effect size",
       title=paste0("MVMR results for ",current_trait_category , ", 95% CI"))+
  theme(legend.position = "none")

ggsave(paste0("figures/mvmr_BMIs-to-", current_trait_category, ".png"),
       plot=p, #scale=1, 
       width=16, height=19,
       #width=10, height=5,
       units=c("cm"), dpi=300, limitsize=F)

```



## MVMR analysis 2: childhood BMI + mediator -> breast cancer

```{r message=F}
mvmr_method<- #"_using_MVMR" # or "" or "_using_2SMR"

dat <- read_tsv( paste0(results_path, current_trait_category, "/merged/merged_mvmr_earlyBMI-", current_trait_category, "-to-BCAC_", bc_data, mvmr_method, ".tsv"))

if (current_trait_category == "hormones"){ dat<-filter(dat, !grepl(",", exposure)) } # don't include testosterone subcats

dat$exposure <- forcats::fct_inorder(dat$exposure) %>% forcats::fct_rev() 
p<-ggplot(dat, aes(y=exposure, x=or, label=outcome, colour=trait_type)) +
  geom_errorbarh(aes(xmin=or_lci95, xmax=or_uci95), height=.3) +
  geom_point(size=2)+
  scale_color_manual(values=pal)+
  geom_vline(xintercept=1, linetype='longdash') +
  theme_minimal_hgrid(10, rel_small = 1) +
  facet_wrap(~outcome, ncol=1)+
  labs(color = "",y = "", x = "OR",
       title=paste0("MVMR results for ",current_trait_category , ", 95% CI"))+
  theme(legend.position = "none")

ggsave(paste0("figures/mvmr_earlyBMI-", current_trait_category, "-to-BCAC_", bc_data, mvmr_method, ".png"),
       plot=p, #scale=1, 
       width=15, height=15,
       #width=10, height=5,
       units=c("cm"), dpi=300, limitsize=F)
```



## MVMR analysis 3: childhood BMI + adult BMI + mediator -> breast cancer

```{r message=F}
dat <- read_tsv( paste0(results_path, current_trait_category, "/merged/merged_mvmr_adultBMI-earlyBMI-", current_trait_category, "-to-BCAC_", bc_data,".tsv"))

if (current_trait_category == "hormones"){ dat<-filter(dat, !grepl(",", exposure)) } # don't include testosterone subcats

# swap second (mediator) and third (Adult BMI) rows in df: 
# this will allow keeping BMIs on top of the viz
dat<-bind_rows(dat[c(1,3,2),], dat[c(4:dim(dat)[1]),])

dat$exposure <- forcats::fct_inorder(dat$exposure) %>% forcats::fct_rev() 
p<-ggplot(dat, aes(y=exposure, x=or, label=outcome, colour=trait_type)) +
  geom_errorbarh(aes(xmin=or_lci95, xmax=or_uci95), height=.3) +
  geom_point(size=2)+
  scale_color_manual(values=pal)+
  geom_vline(xintercept=1, linetype='longdash') +
  theme_minimal_hgrid(10, rel_small = 1) +
  facet_wrap(~outcome, ncol=1)+
  labs(color = "",y = "", x = "OR",
       title=paste0("MVMR results for ",current_trait_category , ", 95% CI"))+
  theme(legend.position = "none")

ggsave(paste0("figures/mvmr_adultBMI-earlyBMI-", current_trait_category, "-to-BCAC_", bc_data,".png"),
       plot=p, #scale=1, 
       width=15, height=16,
       #width=10, height=5,
       units=c("cm"), dpi=300, limitsize=F)
```

## MVMR analysis 4: childhood BMI + childhood height + mediator -> breast cancer

```{r message=F}
mvmr_method<- "_using_2SMR"#"_using_MVMR" # or "" or "_using_2SMR"
bc_data<-"1128"

dat <- read_tsv( paste0(results_path, current_trait_category, "/merged/merged_mvmr_height-earlyBMI-", current_trait_category, "-to-BCAC_", bc_data, mvmr_method,".tsv"))

if (current_trait_category == "hormones"){ dat<-filter(dat, !grepl(",", exposure)) } # don't include testosterone subcats

# swap second (mediator) and third (Adult BMI) rows in df: 
# this will allow keeping BMIs on top of the viz
dat<-bind_rows(dat[c(1,3,2),], dat[c(4:dim(dat)[1]),])

dat$exposure <- forcats::fct_inorder(dat$exposure) %>% forcats::fct_rev() 
p<-ggplot(dat, aes(y=exposure, x=or, label=outcome, colour=trait_type)) +
  geom_errorbarh(aes(xmin=or_lci95, xmax=or_uci95), height=.3) +
  geom_point(size=2)+
  scale_color_manual(values=pal)+
  geom_vline(xintercept=1, linetype='longdash') +
  theme_minimal_hgrid(10, rel_small = 1) +
  facet_wrap(~outcome, ncol=1)+
  labs(color = "",y = "", x = "OR",
       title=paste0("MVMR results for ",current_trait_category , ", 95% CI"))+
  theme(legend.position = "none")

ggsave(paste0("figures/mvmr_height-earlyBMI-", current_trait_category, "-to-BCAC_", bc_data, mvmr_method, ".png"),
       plot=p, #scale=1, 
       width=15, height=16,
       #width=10, height=5,
       units=c("cm"), dpi=300, limitsize=F)
```





# Merging data of all trait together (of analysis2) 
```{r message =F}
mvmr_method<- ""  # "_using_MVMR" # or "" or "_using_2SMR"

# read in  merged files for all traits
trait_categories <- c("hormones", "reproductive_traits", "glycemic_traits")

merged_data <- data.frame()

for (current_trait_category in trait_categories){
  dat<- read_tsv(paste0(results_path,  current_trait_category, 
                        "/merged/merged_mvmr_earlyBMI-", current_trait_category, "-to-BCAC_", bc_data, mvmr_method,".tsv")) 
  dat<-dat %>% 
        mutate(trait_category = stringr::str_c(stringr::str_split(stringr::str_to_title(current_trait_category), "_")[[1]], 
                                      collapse=" ")) 
  merged_data<-bind_rows(merged_data, dat)
}

merged_data<-select(merged_data, tidyselect::any_of( c("exposure", "outcome",
                     "or", "or_lci95", "or_uci95", "pval" ,
                     "nsnp", "trait_category" ,  "trait_type"  )))

mvmr_method<- "_using_2SMR"        
merged_data %>% 
      filter(exposure != "Childhood BMI") %>% 
      tidy_pvals() %>% 
      write_tsv(paste0(results_path, "_merged_results/results_MVMR_mediators_BCAC", bc_data, mvmr_method,".tsv"))

merged_data %>% 
      write_tsv(paste0(results_path, "_merged_results/results_MVMR_all_BCAC", bc_data, mvmr_method, ".tsv"))
```

# Make combined plots of (all) direct estimates

```{r}
mvmr_method<- "_using_2SMR"  
merged_data<-read_tsv(paste0(results_path, "_merged_results/results_MVMR_mediators_BCAC", bc_data, mvmr_method, ".tsv"))

merged_data <- merged_data %>% 
              filter(!grepl("-meno", exposure)) %>% 
              filter(exposure %in% 
                      c("IGF", "Oestradiol" , "SHBG" , "Testosterone (bioavailable)","Testosterone (free)", "Testosterone (total)",
                      "Age at first live birth (UKB)", "Age at menarche (Perry)","Age at menopause (Day)", "Number of births (UKB)",
                      "fasting glucose (Lagou)", "fasting insulin (Lagou)",  "HBa1c (Wheeler)", "HOMA-B", "HOMA-IR",
                      "Childhood height", "Breast size (Ooi 2019)", "Percent mammographic density  (Brand 2018)", "Mammographic density (dense area) (Brand 2018)")) %>% 
              mutate(exposure = ifelse(trait_category != "Hormones", 
                                       gsub(" \\(.*\\)", "", exposure), exposure)) %>% 
              mutate(exposure = case_when(exposure == "fasting insulin" ~ "Fasting Insulin",
                                         exposure == "fasting glucose" ~ "Fasting Glucose",
                                         exposure == "IGF" ~ "IGF-1",
                                         TRUE ~ exposure))
              
  
# subset only to prioritised data
merged_data <- merged_data %>% 
              filter(trait_type %in% c("IGF", "SHBG","Testosterone", "menarche", "menopause", "HBa1c") )


merged_data$exposure <- factor(merged_data$exposure , levels = sort(merged_data$exposure))
merged_data$exposure<- forcats::fct_rev(merged_data$exposure)

merged_data$trait_category <- forcats::fct_inorder(merged_data$trait_category)
#write_tsv(merged_data, paste0(results_path, "_merged_results/results_MVMR_mediators_BCAC", bc_data, mvmr_method,"_fig_input.tsv"))

pal<-c(wes_palette("Darjeeling1")[c(2:5)])
p<-ggplot(merged_data, aes(y=exposure, x=or, label=outcome , group = trait_category, colour=trait_category)) +
  geom_errorbarh(aes(xmin=or_lci95, xmax=or_uci95), height=.2) +
  geom_point(size=2, shape=20)+
  scale_color_manual(values=pal)+
  #xlim(0.43, 2.5)+
  geom_vline(xintercept=1, linetype='longdash') +
  theme_minimal_grid(9) +
  facet_grid(trait_category~outcome, scale="free", switch="x")+
  theme(strip.text = element_text(face = 'bold'))+
  labs(color = "",y = "", x = "Odds ratio", subtitle="",
       title= paste0("                   MVMR: direct effect of mediators on breast cancer (OR, 95% CI)") )+
  theme(legend.position = "none",  plot.title.position = "plot")


ggsave(paste0("figures/merged_mediators_direct-to-BCAC_", bc_data,"_IVW", mvmr_method,".png"),
       plot=p, scale=1.2, 
       width=14, height=12,
       units=c("cm"), dpi=300, limitsize=F)
```


##  show mediators' effect on childhood BMI 

```{r}

dat <- read_tsv(paste0(results_path, "_merged_results/results_MVMR_all_BCAC", bc_data, mvmr_method, ".tsv"))
# add pair labels
dat$pair <- rep(dat$exposure[c(FALSE, TRUE)], each=2)
# don't include testosterone subcats
dat <- filter(dat, !grepl("-meno", pair))  


# OPTION 1: load BMIs-BC mvmr results
basic_model<-fread(paste0(results_path,"bmi/merged/mvmr_BMIs-BCAC_", bc_data, mvmr_method, ".tsv"))  
basic_model$pair <- "Adult BMI"
# OPTION 2: load univariate BMIs-BC results
basic_model2<-fread(paste0(results_path,"bmi/merged/merged_mr_mediators_bmi-to-BCAC_", bc_data,".tsv"))  %>% filter(method == "Inverse variance weighted") 
basic_model2$pair <- "nothing"

# add basic model to dat
dat <- bind_rows(basic_model, basic_model2 ,dat) %>%
      select(-method, -id.outcome, -b, -se, -lo_ci, -up_ci) %>% 
      mutate(trait_category = ifelse(pair %in% c("Adult BMI", "nothing"), "BMI",trait_category ))
# tmp save for making tables
dat %>% filter(exposure == "Childhood BMI") %>% write_tsv(paste0(results_path, "_merged_results/results_MVMR_cBMI_BCAC", bc_data, mvmr_method, ".tsv"))

dat <- dat %>% 
 filter(pair %in% c( "nothing", "Adult BMI",
                    "IGF", "Oestradiol" , "SHBG" , "Testosterone (bioavailable)","Testosterone (free)", "Testosterone (total)",
                      "Age at first live birth (UKB)", "Age at menarche (Perry)","Age at menopause (Day)", "Number of births (UKB)",
                      "fasting glucose (Lagou)", "fasting insulin (Lagou)",  "HBa1c (Wheeler)", "HOMA-B", "HOMA-IR",
                      "Childhood height", "Breast size (Ooi 2019)", "Percent mammographic density  (Brand 2018)", "Mammographic density (dense area) (Brand 2018)")) %>% 
              mutate(pair = ifelse(trait_category != "Hormones", 
                                       gsub(" \\(.*\\)", "", pair), pair)) %>%
              mutate(pair = case_when(pair == "fasting insulin" ~ "Fasting Insulin",
                                         pair == "fasting glucose" ~ "Fasting Glucose",
                                         pair == "IGF" ~ "IGF-1",
                                         pair == "Childhood height" ~ "Age 10 height",
                                         TRUE ~ pair)) 
# subset only to prioritised data
dat <- dat %>% filter(trait_type %in% c("IGF", "SHBG","Testosterone", "menarche", "menopause", "HBa1c", NA) )



# create a new label
dat$direct <- paste0("Childhood body size accounted for " , dat$pair)
# create other custom label
dat$nsnp_mediator <- rep(dat$nsnp[c(FALSE, TRUE)], each=2)
dat$custom_label <- paste0(dat$pair, " (nsnp cBMI: ", dat$nsnp, "; nsnp Med: ", dat$nsnp_mediator, ")" )


# add text colour variable
dat <- mutate( dat, text_col = ifelse(pair %in% c("Adult BMI", "nothing"), "steelblue", 
                               ifelse (nsnp_mediator < 10 | is.na(nsnp_mediator), "grey60",
                                       "black")))
# BMI ot no BMI
dat <- mutate( dat, bmi_no_bmi = ifelse(trait_category == "BMI", "Yes", "No"))

# drop mediator measures
dat_BMI<- dat %>% filter(exposure == "Childhood BMI") %>%  arrange(-or)
dat_BMI %>% kable_it()


# BMI
cBMI_OR <- basic_model %>% filter(exposure == "Childhood BMI") %>% pull(or)
cBMI_OR2 <- basic_model2 %>% filter(exposure == "Childhood BMI") %>% pull(or)

#pal<-c(wes_palette("Darjeeling1")[c(1,4,2,5,3)])
dat_BMI$outcome<-"Breast cancer"
dat_BMI<- dat_BMI %>% mutate(direct = gsub("BMI", "body size", direct)) %>% 
                      mutate(direct = gsub("accounted for nothing", "(total effect)", direct))
dat_BMI <- mutate( dat_BMI, effect = ifelse(grepl('total effect', direct), "total", "direct"))
 
p2<-ggplot(dat_BMI, aes(y=reorder(direct, -or), x=or, label=outcome, colour=effect)) +
  geom_errorbarh(aes(xmin=or_lci95, xmax=or_uci95), height=.3) +
  geom_point(size=2)+
  geom_vline(xintercept=cBMI_OR2, linetype=2, colour = "steelblue") +
  geom_vline(xintercept=1, linetype='longdash') +
  theme_minimal_hgrid(10) +
  scale_color_manual(values=c("tomato", "steelblue"))+
  scale_y_discrete(position = "right")+
  theme(strip.text = element_text(face = 'bold'))+
  facet_grid(~outcome, scale="free", switch="x")+
  labs(color = "",y = "", x = "Odds ratio", subtitle="")+
  theme(legend.position = "none")+
  theme(axis.text.y = element_text(colour = dat_BMI$text_col))

dat_BMI %>%  write_tsv(paste0("figures/plotting_data/fig4.tsv"))

ggsave(paste0("figures/merged_cBMI_direct-to-BCAC_", bc_data,"_IVW", mvmr_method,"_subset_tidy3.pdf"),
       plot=p2, scale=1.5, bg='white',
       width=10, height=6,
       units=c("cm"), dpi=300, limitsize=F)




## Get just the mediators

#dat <- mutate(dat, pval = scales::scientific(pval, digits = 2))
#dat$mediator_label <- paste0(dat$exposure, " (nsnp: ", dat$nsnp, "; pval: ", dat$pval, ")" )
## drop mediator measures
#dat_med<- dat %>% filter(!exposure %in% c("Adult BMI", "Childhood BMI")) %>%  arrange(-or)
#dat_med %>% kable_it()
## mediators
#p2<-ggplot(dat_med, aes(y=reorder(mediator_label, -or), x=or, label=outcome, colour=exposure)) +
#  geom_errorbarh(aes(xmin=or_lci95, xmax=or_uci95), height=.3) +
#  geom_point(size=2)+
#  geom_vline(xintercept=1, linetype='longdash') +
#  theme_minimal_hgrid(10, rel_small = 1) +
#  facet_wrap(~outcome, ncol=1)+
#  labs(color = "",y = "", x = "OR",
#       title="               MVMR results: Mediators accounted for childhood BMI, 95% CI")+
#  theme(legend.position = "none", plot.title.position = "plot")+
#  theme(axis.text.y = element_text(colour = dat_med$text_col))
```



# Merging data of all traits together (of analysis3) 
```{r message =F}
mvmr_method<- ""  # "_using_MVMR" # or "" or "_using_2SMR"

# read in  merged files for all traits
trait_categories <- c("hormones", "reproductive_traits", "glycemic_traits")

merged_data <- data.frame()

for (current_trait_category in trait_categories){
  dat<- read_tsv(paste0(results_path,  current_trait_category, 
                        "/merged/merged_mvmr_adultBMI-earlyBMI-", current_trait_category, "-to-BCAC_", bc_data, mvmr_method,".tsv")) 
  dat<-dat %>% 
        mutate(trait_category = stringr::str_c(stringr::str_split(stringr::str_to_title(current_trait_category), "_")[[1]], 
                                      collapse=" ")) 
  merged_data<-bind_rows(merged_data, dat)
}

merged_data<-select(merged_data, tidyselect::any_of( c("exposure", "outcome",
                     "or", "or_lci95", "or_uci95", "pval" ,
                     "nsnp", "trait_category" ,  "trait_type"  )))

mvmr_method<- "_using_2SMR"  
merged_data %>% 
      filter(!exposure %in% c("Adult BMI", "Childhood BMI")) %>% 
      tidy_pvals() %>%
      write_tsv(paste0(results_path, "_merged_results/results_MVMR_mediators_adultBMI_BCAC", bc_data, mvmr_method,".tsv"))

merged_data %>% 
      write_tsv(paste0(results_path, "_merged_results/results_MVMR_adultBMI_all_BCAC", bc_data, mvmr_method, ".tsv"))
```



## Show all effect on childhood BMI, also accounting for Adult BMI

```{r}
# read in  merged files for all traits
all_categories <- c("hormones", "reproductive_traits", "glycemic_traits", "physical_traits")
merged_files <-c()
for (current_trait_category in all_categories){
  merged_files <- c(merged_files, 
                    paste0(results_path, current_trait_category, "/merged/merged_mvmr_adultBMI-earlyBMI-", current_trait_category, "-to-BCAC_", bc_data,".tsv"))
}

l <- lapply(merged_files, fread)
dat <- l %>% purrr::reduce(bind_rows) 

# add pair labels
dat$pair <- dat %>% filter(!exposure %in% c("Adult BMI", "Childhood BMI")) %>% pull(exposure) %>%  rep(., each=3)
dat <- filter(dat, !grepl(",", pair))  # don't include testosterone subcats

# add nsnp of mediator used to run each mvmr
dat$nsnp_mediator <- dat %>% filter(!exposure %in% c("Adult BMI", "Childhood BMI")) %>% pull(nsnp) %>%  rep(., each=3)

# load BMIs-BC mvmr results
basic_model<-fread(paste0(results_path,"bmi/merged/mvmr_BMIs-BCAC_", bc_data,".tsv"))
basic_model$pair <- "only Adult BMI"
basic_model$nsnp_mediator <- NA

# add basic model to dat
dat <- bind_rows(basic_model, dat)

# create a new label
dat$direct <- paste0("Childhood BMI accounted for Adult BMI and " , dat$pair)

dat$custom_label <- paste0(dat$pair, " (nsnp cBMI: ", dat$nsnp, "; nsnp Med: ", dat$nsnp_mediator, ")" )


# add text colour variable
dat <- mutate( dat, text_col = ifelse(pair =="only Adult BMI", "blue", 
                               ifelse (nsnp_mediator < 10 , "grey", "black")))



# drop mediator measures
dat_BMI<- dat %>% filter(exposure == "Childhood BMI") %>%  arrange(-or)
dat_BMI %>% kable_it()

# BMI
cBMI_OR <- basic_model %>% filter(exposure == "Childhood BMI") %>% pull(or)
p3<-ggplot(dat_BMI, aes(y=reorder(custom_label, -or), x=or, label=outcome, colour=exposure)) +
  geom_errorbarh(aes(xmin=or_lci95, xmax=or_uci95), height=.3) +
  geom_point(size=2)+
  geom_vline(xintercept=cBMI_OR, linetype=2, colour = "blue") +
  geom_vline(xintercept=1, linetype='longdash') +
  theme_minimal_hgrid(10, rel_small = 1) +
  facet_wrap(~outcome, ncol=1)+
  labs(color = "",y = "", x = "OR",
       title="            MVMR results: Childhood BMI accounted for Adult BMI and ... ")+
  theme(legend.position = "none", plot.title.position = "plot")+
  theme(axis.text.y = element_text(colour = dat_BMI$text_col))


## Get just the mediators

dat <- mutate(dat, pval = scales::scientific(pval, digits = 2))
dat$mediator_label <- paste0(dat$exposure, " (nsnp: ", dat$nsnp, "; pval: ", dat$pval, ")" )
# drop mediator measures
dat_med<- dat %>% filter(!exposure %in% c("Childhood BMI", "Adult BMI")) %>%  arrange(-or)

# mediators
p4<-ggplot(dat_med, aes(y=reorder(mediator_label, -or), x=or, label=outcome, colour=exposure)) +
  geom_errorbarh(aes(xmin=or_lci95, xmax=or_uci95), height=.3)+
  geom_point(size=2)+
  geom_vline(xintercept=1, linetype='longdash') +
  theme_minimal_hgrid(10, rel_small = 1) +
  facet_wrap(~outcome, ncol=1)+
  labs(color = "",y = "", x = "OR",
       title="       MVMR results: Mediators accounted for childhood BMI and adult BMI, 95% CI")+
  theme(legend.position = "none", plot.title.position = "plot")+
  theme(axis.text.y = element_text(colour = dat_med$text_col))

```



# Merging data of all traits together (of analysis4) 
```{r message =F}
mvmr_method<- "_using_2SMR"  # "_using_MVMR" # or "" or "_using_2SMR"

# read in  merged files for all traits
trait_categories <- c("hormones", "reproductive_traits", "glycemic_traits")

merged_data <- data.frame()

for (current_trait_category in trait_categories){
  dat<- read_tsv(paste0(results_path,  current_trait_category, 
                        "/merged/merged_mvmr_height-earlyBMI-", current_trait_category, "-to-BCAC_", bc_data, mvmr_method,".tsv")) 
  dat<-dat %>% 
        mutate(trait_category = stringr::str_c(stringr::str_split(stringr::str_to_title(current_trait_category), "_")[[1]], 
                                      collapse=" ")) 
  merged_data<-bind_rows(merged_data, dat)
}

merged_data<-select(merged_data, tidyselect::any_of( c("exposure", "outcome",
                     "or", "or_lci95", "or_uci95", "pval" ,
                     "nsnp", "trait_category" ,  "trait_type"  )))
            
merged_data %>% 
      filter(!exposure %in% c("Childhood height", "Childhood BMI")) %>% 
      tidy_pvals() %>% 
      write_tsv(paste0(results_path, "_merged_results/results_MVMR_mediators_height_BCAC", bc_data, mvmr_method,".tsv"))

merged_data %>% 
      write_tsv(paste0(results_path, "_merged_results/results_MVMR_height_all_BCAC", bc_data, mvmr_method, ".tsv"))
```
	
	
## new analysis: try putting together estimates from MVMR (+adult BMI, childhood heigth)

```{r}
mvmr_method<- "_using_2SMR"  

# MVMR cBMI + mediator
data1<-read_tsv(paste0(results_path, "_merged_results/results_MVMR_mediators_BCAC", bc_data, mvmr_method, ".tsv")) %>% 
       select(exposure, or, or_lci95, or_uci95) %>% 
       rename('or_data1' = 'or', "or_lci95_data1" = "or_lci95", "or_uci95_data1"="or_uci95")

# MVMR cBMI + mediator + aBMI
data2<-read_tsv(paste0(results_path, "_merged_results/results_MVMR_mediators_adultBMI_BCAC", bc_data, mvmr_method,".tsv")) %>% 
       select(exposure, or, or_lci95, or_uci95) %>% 
       rename('or_data2' = 'or', "or_lci95_data2" = "or_lci95", "or_uci95_data2"="or_uci95")


# MVMR cBMI + mediator + childhood height
data3<-read_tsv(paste0(results_path, "_merged_results/results_MVMR_mediators_height_BCAC", bc_data, mvmr_method,".tsv")) %>% 
        select(exposure, or, or_lci95, or_uci95) %>% 
       rename('or_data3' = 'or', "or_lci95_data3" = "or_lci95", "or_uci95_data3"="or_uci95")


# join and tidy up data
data_all <- full_join(data1, data2, by="exposure") %>% full_join(., data3, by="exposure") %>% 
              filter(!grepl("-meno", exposure)) %>% 
              drop_na()

# convert to format for plotting  
data_all_reshaped<-
  data_all %>% 
    pivot_longer(cols = c( "or_data1" , "or_data2" , "or_data3"), names_to = "or", values_to = "indirect_or") %>%
    pivot_longer(cols = c( "or_lci95_data1" , "or_lci95_data2" , "or_lci95_data3"), names_to = "or_lci95", values_to = "indirect_or_lci95") %>% 
    pivot_longer(cols = c( "or_uci95_data1" , "or_uci95_data2" , "or_uci95_data3"), names_to = "or_uci95", values_to = "indirect_or_uci95") %>% 
    mutate(or = gsub("or_", "", or)) %>% 
    mutate(or_lci95 = gsub("or_lci95_", "", or_lci95)) %>%
    mutate(or_uci95 = gsub("or_uci95_", "", or_uci95)) %>% 
    mutate(method = ifelse(or == or_lci95 & or_lci95 == or_uci95, 1, 0)) %>% 
    filter(method == 1) %>% 
    select(-or_lci95, -or_uci95, -method) %>% 
    rename("analysis_type" = "or") %>% 
    mutate(analysis_type= case_when(analysis_type == "data1" ~ "MVMR: childhood body size + mediator",
                                    analysis_type == "data2" ~ "MVMR: childhood body size + adult body size + mediator",
                                    analysis_type == "data3" ~ "MVMR: childhood body size + childhood height + mediator", 
                                    TRUE ~ analysis_type))
  
# set the order in which they apper in th eplot
data_all_reshaped$analysis_type <- factor(data_all_reshaped$analysis_type ,
                               levels = rev(c("MVMR: childhood body size + mediator", 
                                              "MVMR: childhood body size + adult body size + mediator",
                                              "MVMR: childhood body size + childhood height + mediator")))

data_all_reshaped_sub <- data_all_reshaped %>% filter(exposure %in% c("IGF", "SHBG" , "Testosterone (bioavailable)","Testosterone (free)", "Testosterone (total)", "Age at menarche (Perry)","Age at menopause (Day)"))

# set the order in which they apper in th eplot
data_all_reshaped_sub$exposure <- factor(data_all_reshaped_sub$exposure ,
                               levels = rev(c("IGF", "SHBG", "Testosterone (bioavailable)", "Testosterone (free)",  "Testosterone (total)" ,  "Age at menarche (Perry)", "Age at menopause (Day)"   )))


p<-ggplot(data_all_reshaped_sub,
           aes(y=exposure, x=indirect_or, 
               label=analysis_type , group = analysis_type, colour=analysis_type)) +
  geom_point(size=3, shape=20, 
             position = ggstance::position_dodgev(height = 0.5))+
  geom_errorbarh(aes(xmin=indirect_or_lci95, xmax=indirect_or_uci95), height=.1, 
            position = ggstance::position_dodgev(height = 0.5)) +
  scale_x_continuous(breaks=seq(0.8,1.3,0.05))+
  theme_minimal_hgrid(10) +
  geom_vline(xintercept=1, linetype='longdash') +
  scale_color_manual(values=c("#FF0060", "#B80041", "#5E0224"))+
  theme(strip.text = element_text(face = 'bold'),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(),
        )+
  labs(color = "",y = "", x = "OR of direct effect"
       #subtitle="                   accounted for adult BMI or childhood height: ",
       #title= "               MVMR: direct effect of mediators on breast cancer" 
       )+
  theme(legend.position = "bottom", legend.direction = "vertical", 
        plot.title.position = "plot")
```
## Putting total and direct effects of mediators together in one plot
```{r}
data1 <- 
  read_tsv(paste0(results_path, "_merged_results/results_MR_mediators_BCAC", bc_data,"_fig_input.tsv")) %>% 
  select(exposure, or, or_lci95, or_uci95) %>% 
  rename('or_data1' = 'or', "or_lci95_data1" = "or_lci95", "or_uci95_data1"="or_uci95")
data2 <- 
  read_tsv(paste0(results_path, "_merged_results/results_MVMR_mediators_BCAC", bc_data, mvmr_method,"_fig_input.tsv")) %>% 
  select(exposure, or, or_lci95, or_uci95) %>% 
  rename('or_data2' = 'or', "or_lci95_data2" = "or_lci95", "or_uci95_data2"="or_uci95")
# join and tidy up data
data_all <- full_join(data1, data2, by="exposure") %>% 
              drop_na()
# convert to format for plotting  
data_all_reshaped<-
  data_all %>% 
    pivot_longer(cols = c( "or_data1" , "or_data2" ), names_to = "or", values_to = "indirect_or") %>%
    pivot_longer(cols = c( "or_lci95_data1" , "or_lci95_data2" ), names_to = "or_lci95", values_to = "indirect_or_lci95") %>% 
    pivot_longer(cols = c( "or_uci95_data1" , "or_uci95_data2" ), names_to = "or_uci95", values_to = "indirect_or_uci95") %>% 
    mutate(or = gsub("or_", "", or)) %>% 
    mutate(or_lci95 = gsub("or_lci95_", "", or_lci95)) %>%
    mutate(or_uci95 = gsub("or_uci95_", "", or_uci95)) %>% 
    mutate(method = ifelse(or == or_lci95 & or_lci95 == or_uci95, 1, 0)) %>% 
    filter(method == 1) %>% 
    select(-or_lci95, -or_uci95, -method) %>% 
    rename("analysis_type" = "or") %>% 
    mutate(analysis_type= case_when(analysis_type == "data1" ~ "Total effect (uniMR)",
                                    analysis_type == "data2" ~ "Direct effect (MVMR)",
                                    TRUE ~ analysis_type))
  
# set the order in which they apper in th eplot
data_all_reshaped$analysis_type <- factor(data_all_reshaped$analysis_type ,
                               levels = rev(c("Direct effect (MVMR)",
                                              "Total effect (uniMR)")))
data_all_reshaped_sub <-
  data_all_reshaped %>% 
  filter(exposure %in% c("IGF-1", "SHBG" , "Testosterone (bioavailable)","Testosterone (free)", 
                         "Testosterone (total)", "Age at menarche","Age at menopause", "HBa1c"))
# set the order in which they apper in th eplot
data_all_reshaped_sub$exposure <- 
                  factor(data_all_reshaped_sub$exposure ,
                          levels = rev(c("IGF-1", "SHBG" , "Testosterone (bioavailable)","Testosterone (free)", 
                         "Testosterone (total)", "Age at menarche","Age at menopause" , "HBa1c")))
p<-ggplot(data_all_reshaped_sub,
           aes(y=exposure, x=indirect_or, 
               label=analysis_type , group = analysis_type, colour=analysis_type, shape=analysis_type)) +
  geom_point(size=3, shape=20, 
             position = ggstance::position_dodgev(height = 0.5))+
  geom_errorbarh(aes(xmin=indirect_or_lci95, xmax=indirect_or_uci95), height=.1, 
            position = ggstance::position_dodgev(height = 0.5)) +
  scale_x_continuous(breaks=seq(0.7,1.4,0.1))+
  theme_minimal_hgrid(10) +
  geom_vline(xintercept=1, linetype='longdash') +
  scale_shape_manual(values = c(19,17)) +
  scale_color_manual(values=c("#0E5888", "#9BD7D6"))+
  theme(strip.text = element_text(face = 'bold'),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line())+
  labs(color = "",y = "", x = "OR"
       #subtitle="                   from step2 MR and MVMR analyses ",
       #title= "         MR: total and direct effects of mediators on breast cancer (OR, 95%)" 
       )+
  theme(legend.position = "bottom", legend.direction = "vertical", 
        plot.title.position = "plot")
```












## BMI plots only 

```{r}
current_trait_category <- "bmi"
dat_univ<- read_tsv(paste0(results_path, current_trait_category, 
                           "/merged/merged_mr_mediators_",current_trait_category,"-to-BCAC_", bc_data,".tsv")) %>%
  filter(method %in% c("Inverse variance weighted", "Wald ratio")) %>% 
  select(-id.outcome, -method) %>% 
  mutate(mr="Univariable MR")


dat_mv <- read_tsv(paste0(results_path, current_trait_category, "/merged/mvmr_BMIs-BCAC_", bc_data,"_using_2SMR.tsv")) %>% 
  mutate(mr="Multivariable MR")

dat<-bind_rows(dat_univ, dat_mv)

dat %>% kable_it()

pal <- wes_palette("Zissou1")[c(1,5)]
dat$mr <- forcats::fct_rev(dat$mr) 
p<-ggplot(dat, aes(y=exposure, x=or, label=outcome, colour=mr,  shape=exposure)) +
  geom_errorbarh(aes(xmin=or_lci95, xmax=or_uci95), height=.3) +
  geom_point(size=3)+
  scale_color_manual(values=pal)+
  scale_shape_manual(values = c(16,20)) +
  geom_vline(xintercept=1, linetype='longdash') +
  theme_minimal_hgrid(10, rel_small = 1) +
  facet_wrap(mr~outcome, ncol=1)+
  labs(color = "",y = "", x = "Odds ratio",
       title= paste0("        MR results for ", toupper(current_trait_category) ," to Breast Cancer, 95% CI") )+
  theme(legend.position = "none", plot.title.position = "plot")

ggsave(paste0("figures/MR_", toupper(current_trait_category) ,"-to_BCAC_", bc_data,"_IVW.png"),
       plot=p, scale=1, 
       width=10, height=7,
       units=c("cm"), dpi=300, limitsize=F)
scales::show_col(c(wes_palette("GrandBudapest2")))
```



