---
title: "generate forest plots"
author: "Marina Vabistsevits"
date: "10/04/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library(tidyr)
library(purrr)
library(readr)
library(tibble)
library(dplyr)
library(data.table)
library(vroom)
library(TwoSampleMR)
library(ggplot2)
library(cowplot)
library(wesanderson)
```

```{r}
# set path for pre-calculated data, outside the code repo
# `local` / `remote` (reading data from RDSF)
currently_working_env = "local"
source("set_paths.R")
set_paths(currently_working_env)

# metafile
data_lookup<-read_csv(paste0("metadata/data_lookup.csv")) 
```

## MVMR 2 BMI - mediator
```{r}
all.files_mr <- list.files(path = results_path, pattern = "mvmr_BMIs-*", full.names = T)
l <- lapply(all.files_mr, fread)

df <- l %>% reduce(rbind) 

write_tsv(df, paste0(results_path, "mvmr_BMIs-mediators_using_2SMR.tsv"))
df<-read_tsv(paste0(results_path, "mvmr_BMIs-mediators_using_2SMR.tsv"))


pal <- wes_palette("Darjeeling1")[c(2,3,4,5)]
p<-ggplot(df, aes(y=exposure, x=b, label=outcome.full, shape=exposure, colour=outcome)) +
  geom_errorbarh(aes(xmin=lo_ci, xmax=up_ci), height=.3) +
  geom_point(size=2)+##, shape=c(20,16))+#, color = "orange") +
  scale_color_manual(values=pal)+
  scale_shape_manual(values = c(19,20)) +
  #coord_fixed(ratio=.3) +
  geom_vline(xintercept=0, linetype='longdash') +
  theme_minimal_hgrid(10, rel_small = 1) +
  facet_wrap(~outcome.full, ncol=1)+
  labs(color = "",y = "", x = "Effect size",
       title="MVMR results, 95% CI")+
  theme(legend.position = "none")


```


## MVMR early BMI + mediator -> BC
```{r}
all.files_mr <- list.files(path = results_path, pattern = "mvmr_earlyBMI-*", full.names = T)
l <- lapply(all.files_mr[c(11,12,13,20,21)], fread)

df <- l %>% reduce(rbind)


df$exposure.pair<-c("IGF"         ,"IGF" ,
                    "Oestradiol"  ,"Oestradiol",
                    "SHBG"        ,"SHBG"  ,
                    "SHBG"        ,"SHBG"  ,
                    "Testosterone","Testosterone",
                    "Testosterone","Testosterone",
                    "Testosterone","Testosterone",
                    "Testosterone","Testosterone",
                    "Testosterone","Testosterone")

write_tsv(df, paste0(results_path, "mvmr_earlyBMI_mediators_BC_using_2SMR.tsv"))

df<-read_tsv(paste0(results_path, "mvmr_earlyBMI_mediators_BC_using_2SMR.tsv"))




pal <- wes_palette("Darjeeling1")[c(2,3,4,5)]
p<-ggplot(df, aes(y=exposure, x=or, label=outcome, colour=exposure.pair)) +
  geom_errorbarh(aes(xmin=or_lci95, xmax=or_uci95), height=.3) +
  geom_point(size=2)+##, shape=c(20,16))+#, color = "orange") +
  scale_color_manual(values=pal)+
  #scale_shape_manual(values = c(19,20)) +
  #coord_fixed(ratio=.3) +
  geom_vline(xintercept=1, linetype='longdash') +
  theme_minimal_hgrid(10, rel_small = 1) +
  facet_wrap(~outcome, ncol=1)+
  labs(color = "",y = "", x = "Effect size",
       title="MVMR results, 95% CI")+
  theme(legend.position = "none")


```

## MVMR early BMI + testosterones -> BC
```{r}
all.files_mr <- list.files(path = results_path, pattern = "mvmr_earlyBMI-*", full.names = T)
l <- lapply(all.files_mr[grepl("Testosterone", all.files_mr)], fread)

df <- l %>% reduce(rbind)
# adhoc add full names
write_tsv(df, paste0(results_path, "mvmr_earlyBMI_testosterones_BC_using_2SMR.tsv"))

df<-df %>% 
  mutate(measure = ifelse(grepl("free", exposure), "Free", 
                   ifelse(grepl("bioavailable", exposure), "Bioavailable",
                   ifelse(grepl("total", exposure), "Total", "Total (old)")))) %>% 
  mutate(menopause = ifelse(grepl("pre", exposure), "pre-menopause", 
                   ifelse(grepl("post", exposure), "post-menopause", "all"))) %>% 
  mutate(menopause = factor(menopause, levels=c( "pre-menopause", "post-menopause", "all"))) %>% 
  arrange(menopause)

pal<-unname(yarrr::piratepal("google"))
p<-ggplot(df, aes(y=exposure, x=or, label=outcome, colour=measure)) +
  geom_errorbarh(aes(xmin=or_lci95, xmax=or_uci95), height=.3) +
  geom_point(size=2)+##, shape=c(20,16))+#, color = "orange") +
  scale_color_manual(values=pal)+
  #scale_shape_manual(values = c(19,20)) +
  #coord_fixed(ratio=.3) +
  geom_vline(xintercept=1, linetype='longdash') +
  theme_minimal_hgrid(10, rel_small = 1) +
  facet_wrap(~outcome, ncol=1)+
  labs(color = "",y = "", x = "Effect size",
       title="MVMR results, 95% CI")+
  theme(legend.position = "none")


```

## MVMR BMIs + mediator -> BC

```{r}
all.files_mr <- list.files(path = results_path, pattern = "mvmr_adultBMI-*", full.names = T)
l <- lapply(all.files_mr, fread)

df <- l %>% reduce(rbind)
# adhoc add full names
write_tsv(df, paste0(results_path, "mvmr_adultBMI-earlyBMI_mediators_BC_using_2SMR.tsv"))


df$exposure.pair<-c("IGF"         ,"IGF"         ,"IGF" ,
                    "Oestradiol"  ,"Oestradiol"  ,"Oestradiol",
                    "SHBG"        ,"SHBG"        ,"SHBG"  ,
                    "SHBG"        ,"SHBG"        ,"SHBG"  ,
                    "Testosterone","Testosterone","Testosterone",
                    "Testosterone","Testosterone","Testosterone",
                    "Testosterone","Testosterone","Testosterone",
                    "Testosterone","Testosterone","Testosterone",
                    "Testosterone","Testosterone","Testosterone")

pal <- wes_palette("Darjeeling1")[c(2,3,4,5)]
p<-ggplot(df, aes(y=exposure, x=or, label=outcome, colour=exposure.pair)) +
  geom_errorbarh(aes(xmin=or_lci95, xmax=or_uci95), height=.3) +
  geom_point(size=2)+##, shape=c(20,16))+#, color = "orange") +
  scale_color_manual(values=pal)+
  #scale_shape_manual(values = c(19,20)) +
  #coord_fixed(ratio=.3) +
  geom_vline(xintercept=1, linetype='longdash') +
  theme_minimal_hgrid(10, rel_small = 1) +
  facet_wrap(~outcome, ncol=1)+
  labs(color = "",y = "", x = "Effect size",
       title="MVMR results, 95% CI")+
  theme(legend.position = "none")


```



### MVMR reproductive traits

## MVMR 2 BMI - reptrair
```{r}
all.files_mr <- list.files(path = results_path, pattern = "mvmr_BMIs-*", full.names = T)
l <- lapply(all.files_mr[c(1:5,8)], fread)

df <- l %>% reduce(rbind) 

write_tsv(df, paste0(results_path, "mvmr_BMIs-ReproductiveTraits_using_2SMR.tsv"))

dat<-df %>% 
  mutate(trait_type = ifelse(grepl("menarche", outcome.full), "menarche",
                             ifelse(grepl("menopause", outcome.full), "menopause",
                             ifelse(grepl("live birth", outcome.full), "live birth",
                             ifelse(grepl("births", outcome.full), "births", "BMI"
                             ))))) 


pal<-rev(unname(yarrr::piratepal("info2")))
p<-ggplot(dat, aes(y=exposure, x=b, label=outcome.full, shape=exposure, colour=trait_type)) +
  geom_errorbarh(aes(xmin=lo_ci, xmax=up_ci), height=.3) +
  geom_point(size=2)+##, shape=c(20,16))+#, color = "orange") +
  scale_color_manual(values=pal)+
  scale_shape_manual(values = c(19,20)) +
  #coord_fixed(ratio=.3) +
  geom_vline(xintercept=0, linetype='longdash') +
  theme_minimal_hgrid(10, rel_small = 1) +
  facet_wrap(~outcome.full, ncol=1)+
  labs(color = "",y = "", x = "Effect size",
       title="MVMR results, 95% CI")+
  theme(legend.position = "none")


```

## MVMR early BMI + reptraits -> BC
```{r}
all.files_mr <- list.files(path = results_path, pattern = "mvmr_earlyBMI-*", full.names = T)
l <- lapply(all.files_mr[c(4:8,11)], fread)

df <- l %>% reduce(rbind)
# adhoc add full names
write_tsv(df, paste0(results_path, "mvmr_earlyBMI_ReproductiveTraits_BC_using_2SMR.tsv"))

df$trait_type<-c(rep("live birth", 2),
                    rep("menarche", 4),
                    rep("menopause", 4),
                    rep("births", 2))

pal<-rev(unname(yarrr::piratepal("info2")))[c(1:4,6)]
df$exposure <- forcats::fct_inorder(df$exposure) %>% forcats::fct_rev() 
p<-ggplot(df, aes(y=exposure, x=or, label=outcome, colour=trait_type)) +
  geom_errorbarh(aes(xmin=or_lci95, xmax=or_uci95), height=.3) +
  geom_point(size=2)+##, shape=c(20,16))+#, color = "orange") +
  scale_color_manual(values=pal)+
  #scale_shape_manual(values = c(19,20)) +
  #coord_fixed(ratio=.3) +
  geom_vline(xintercept=1, linetype='longdash') +
  theme_minimal_hgrid(10, rel_small = 1) +
  facet_wrap(~outcome, ncol=1)+
  labs(color = "",y = "", x = "Effect size",
       title="MVMR results, 95% CI")+
  theme(legend.position = "none")

scales::show_col(pal)
```


## MVMR BMIs + mediator -> BC

```{r}
all.files_mr <- list.files(path = results_path, pattern = "mvmr_adultBMI-*", full.names = T)
l <- lapply(all.files_mr[c(2:6,8)], fread)

df <- l %>% reduce(rbind)
# adhoc add full names
write_tsv(df, paste0(results_path, "mvmr_adultBMI-earlyBMI_ReproductiveTraits_BC_using_2SMR.tsv"))
df<-read_tsv(paste0(results_path, "mvmr_adultBMI-earlyBMI_ReproductiveTraits_BC_using_2SMR.tsv"))

df$exposure.pair<-c(rep("live birth", 3),
                    rep("menarche", 6),
                    rep("menopause", 6),
                    rep("births", 3))
df$exposure.pair2<-c(rep("live birth", 3),
                    rep("menarche Perry", 3),
                    rep("menarche UKB", 3),
                    rep("menopause Day", 3),
                    rep("menopause UKB", 3),
                    rep("births", 3))

pal<-rev(unname(yarrr::piratepal("info2")))[c(1:4,6,7)]
df$exposure <- forcats::fct_inorder(df$exposure) %>% forcats::fct_rev() 
p<-ggplot(df, aes(y=exposure, x=or, label=outcome, colour=exposure.pair)) +
  geom_errorbarh(aes(xmin=or_lci95, xmax=or_uci95), height=.3) +
  geom_point(size=2)+##, shape=c(20,16))+#, color = "orange") +
  scale_color_manual(values=pal)+
  #scale_shape_manual(values = c(19,20)) +
  #coord_fixed(ratio=.3) +
  geom_vline(xintercept=1, linetype='longdash') +
  theme_minimal_hgrid(10, rel_small = 1) +
  facet_wrap(~outcome, ncol=1)+
  labs(color = "",y = "", x = "Effect size",
       title="MVMR results, 95% CI")+
  theme(legend.position = "none")

p<-ggplot(df, aes(y=exposure, x=or, label=outcome, colour=exposure.pair)) +
  geom_errorbarh(aes(xmin=or_lci95, xmax=or_uci95), height=.3) +
  geom_point(size=2)+##, shape=c(20,16))+#, color = "orange") +
  scale_color_manual(values=pal)+
  #scale_shape_manual(values = c(19,20)) +
  #coord_fixed(ratio=.3) +
  geom_vline(xintercept=1, linetype='longdash') +
  theme_minimal_hgrid(10, rel_small = 1) +
  facet_wrap(exposure.pair2~outcome, ncol =2)+
  labs(color = "",y = "", x = "Effect size",
       title="MVMR results, 95% CI")+
  theme(legend.position = "none")

```




### MVMR glycemic traits

## MVMR 2 BMI - glytrait
```{r}
all.files_mr <- list.files(path = paste0(results_path, "glycemic_traits/"), pattern = "mvmr_BMIs-*", full.names = T)
l <- lapply(all.files_mr, fread)

df <- l %>% reduce(bind_rows) 

dat<-df %>% 
          mutate(trait_type = ifelse(grepl("insulin", outcome.full), "insulin",
                             ifelse(grepl("glucose", outcome.full), "glucose",
                             ifelse(grepl("HOMA-B", outcome.full), "HOMA-B",
                             ifelse(grepl("HOMA-IR", outcome.full), "HOMA-IR",
                             ifelse(grepl("HBa1c", outcome.full), "HBa1c", NA
                              ))))))

write_tsv(dat, paste0(results_path, "glycemic_traits/", "mvmr_BMIs-glycemic_traits_using_2SMR_", Sys.Date(),".tsv"))

pal<-c(unname(yarrr::piratepal("pony")))
p<-ggplot(dat, aes(y=exposure, x=b, label=outcome.full, shape=exposure, colour=trait_type)) +
  geom_errorbarh(aes(xmin=lo_ci, xmax=up_ci), height=.3) +
  geom_point(size=2)+##, shape=c(20,16))+#, color = "orange") +
  scale_color_manual(values=pal)+
  scale_shape_manual(values = c(19,20)) +
  #coord_fixed(ratio=.3) +
  geom_vline(xintercept=0, linetype='longdash') +
  theme_minimal_hgrid(10, rel_small = 1) +
  facet_wrap(~outcome.full, ncol=1)+
  labs(color = "",y = "", x = "Effect size",
       title="MVMR results, 95% CI")+
  theme(legend.position = "none")

kable_it(dat)
```

## MVMR early BMI + glytraits -> BC
```{r}
all.files_mr <- list.files(path = paste0(results_path, "glycemic_traits/"), pattern = "mvmr_earlyBMI-*", full.names = T)
l <- lapply(all.files_mr, fread)

df <- l %>% reduce(rbind)

df$trait_type<-c(rep("glucose", 4),
                    rep("insulin", 4),
                    rep("HBa1c", 2),
                    rep("HOMA-B", 2),
                    rep("HOMA-IR", 2))

write_tsv(df, paste0(results_path, "glycemic_traits/" ,"mvmr_earlyBMI_glycemic_traits_BC_using_2SMR_", Sys.Date(),".tsv"))

pal<-c(unname(yarrr::piratepal("pony")))
df$exposure <- forcats::fct_inorder(df$exposure) %>% forcats::fct_rev() 
p<-ggplot(df, aes(y=exposure, x=or, label=outcome, colour=trait_type)) +
  geom_errorbarh(aes(xmin=or_lci95, xmax=or_uci95), height=.3) +
  geom_point(size=2)+##, shape=c(20,16))+#, color = "orange") +
  scale_color_manual(values=pal)+
  #scale_shape_manual(values = c(19,20)) +
  #coord_fixed(ratio=.3) +
  geom_vline(xintercept=1, linetype='longdash') +
  theme_minimal_hgrid(10, rel_small = 1) +
  facet_wrap(~outcome, ncol=1)+
  labs(color = "",y = "", x = "OR",
       title="MVMR results, 95% CI")+
  theme(legend.position = "none")

kable_it(df)
```


## MVMR BMIs + mediator -> BC

```{r}
all.files_mr <- list.files(path = paste0(results_path, "glycemic_traits/"), pattern = "mvmr_adultBMI-*", full.names = T)
l <- lapply(all.files_mr, fread)

df <- l %>% reduce(rbind)

df$trait_type<-c(rep("glucose", 6),
                    rep("insulin", 6),
                    rep("HBa1c", 3),
                    rep("HOMA-B", 3),
                    rep("HOMA-IR", 3))

write_tsv(df, paste0(results_path, "glycemic_traits/" , "mvmr_adultBMI-earlyBMI_glycemic_traits_BC_using_2SMR_", Sys.Date(),".tsv"))



pal<-c(unname(yarrr::piratepal("pony")))
df$exposure <- forcats::fct_inorder(df$exposure) %>% forcats::fct_rev() 
p<-ggplot(df, aes(y=exposure, x=or, label=outcome, colour=trait_type)) +
  geom_errorbarh(aes(xmin=or_lci95, xmax=or_uci95), height=.3) +
  geom_point(size=2)+##, shape=c(20,16))+#, color = "orange") +
  scale_color_manual(values=pal)+
  #scale_shape_manual(values = c(19,20)) +
  #coord_fixed(ratio=.3) +
  geom_vline(xintercept=1, linetype='longdash') +
  theme_minimal_hgrid(10, rel_small = 1) +
  facet_wrap(~outcome, ncol=1)+
  labs(color = "",y = "", x = "OR",
       title="MVMR results, 95% CI")+
  theme(legend.position = "none")

kable_it(df)

```





```{r}
kable_it<-function(df){
  library(kableExtra)
  df %>% 
    mutate(pval= as.character(pval)) %>% 
  mutate_if(is.numeric, round, digits=2) %>% 
  mutate(pval=as.numeric(pval),
         pval=scales::scientific(pval, digits = 2)) %>% 
  kable(.) %>%
  kable_styling()
}
```




# Compare the effects on childhood BMI in all mvmr models

```{r}

mvmr_summary_files<-c("mvmr_BMI-to-BC_using_2SMR.tsv",
                      "mvmr_earlyBMI_mediators_BC_using_2SMR.tsv",
                      "reproductive_traits/mvmr_earlyBMI_ReproductiveTraits_BC_using_2SMR.tsv",
                      "mvmr_earlyBMI_testosterones_BC_using_2SMR.tsv",
                      "glycemic_traits/mvmr_earlyBMI_glycemic_traits_BC_using_2SMR_2020-05-18.tsv")

l <- lapply(paste0(results_path, mvmr_summary_files), fread)

df <- l %>% reduce(bind_rows) %>% distinct()

# add pair labels
df$pair <- rep(df$exposure[c(FALSE, TRUE)], each=2)
# create a new label
df$direct <- paste0("Childhood BMI acc. for " , df$pair)

# drop mediator measures
df_BMI<- df %>% filter(exposure == "Childhood BMI") %>%  arrange(or)
df_BMI %>% kable_it()

# BMI
p<-ggplot(df_BMI, aes(y=reorder(direct, -or), x=or, label=outcome, colour=exposure)) +
  geom_errorbarh(aes(xmin=or_lci95, xmax=or_uci95), height=.3) +
  geom_point(size=2)+
  geom_vline(xintercept=0.5855992, linetype=2, colour = "blue") +
  geom_vline(xintercept=1, linetype='longdash') +
  theme_minimal_hgrid(10, rel_small = 1) +
  facet_wrap(~outcome, ncol=1)+
  labs(color = "",y = "", x = "OR",
       title="            MVMR results: Childhood BMI with mediators, 95% CI")+
  theme(legend.position = "none", plot.title.position = "plot")

# drop mediator measures
df_med<- df %>% filter(exposure != "Childhood BMI") %>%  arrange(or)
df_med %>% kable_it()
# mediators
p2<-ggplot(df_med, aes(y=reorder(exposure, -or), x=or, label=outcome, colour=exposure)) +
  geom_errorbarh(aes(xmin=or_lci95, xmax=or_uci95), height=.3) +
  geom_point(size=2)+
  #geom_vline(xintercept=0.5855992, linetype=2, colour = "blue") +
  geom_vline(xintercept=1, linetype='longdash') +
  theme_minimal_hgrid(10, rel_small = 1) +
  facet_wrap(~outcome, ncol=1)+
  labs(color = "",y = "", x = "OR",
       title="               MVMR results: mediator with childhood BMI, 95% CI")+
  theme(legend.position = "none", plot.title.position = "plot")
```

## + adult BMI in the model 
```{r}

mvmr_summary_files<-c("mvmr_adultBMI-earlyBMI_mediators_BC_using_2SMR.tsv",
                       "reproductive_traits/mvmr_adultBMI-earlyBMI_ReproductiveTraits_BC_using_2SMR.tsv",
                       "glycemic_traits/mvmr_adultBMI-earlyBMI_glycemic_traits_BC_using_2SMR_2020-05-18.tsv")

l <- lapply(paste0(results_path, mvmr_summary_files), fread)

df <- l %>% reduce(bind_rows) %>% distinct()

# add pair labels
df$pair <- df %>% filter(!exposure %in% c("Adult BMI", "Childhood BMI")) %>% pull(exposure) %>%  rep(., each=3)

basic_model<-fread(paste0(results_path,"mvmr_BMI-to-BC_using_2SMR.tsv"))
basic_model$pair <- "only Adult BMI"

df<-bind_rows(basic_model, df)

# create a new label
df$direct <- paste0("Childhood BMI acc. for Adult BMI and " , df$pair )

# drop mediator measures
df_BMI<- df %>% filter(exposure == "Childhood BMI") %>%  arrange(or)


df_BMI %>% kable_it()

# BMI
p3<-ggplot(df_BMI, aes(y=reorder(direct, -or), x=or, label=outcome, colour=exposure)) +
  geom_errorbarh(aes(xmin=or_lci95, xmax=or_uci95), height=.3) +
  geom_point(size=2)+
  geom_vline(xintercept=0.5855992, linetype=2, colour = "blue") +
  geom_vline(xintercept=1, linetype='longdash') +
  theme_minimal_hgrid(10, rel_small = 1) +
  facet_wrap(~outcome, ncol=1)+
  labs(color = "",y = "", x = "OR",
       title="            MVMR results: Childhood BMI with Adult BMI mediators, 95% CI")+
  theme(legend.position = "none", plot.title.position = "plot")


# drop mediator measures
df_med<- df %>% filter(!exposure %in% c("Childhood BMI", "Adult BMI")) %>%  arrange(or)

# mediators
p4<-ggplot(df_med, aes(y=reorder(exposure, -or), x=or, label=outcome, colour=exposure)) +
  geom_errorbarh(aes(xmin=or_lci95, xmax=or_uci95), height=.3) +
  geom_point(size=2)+
  #geom_vline(xintercept=0.5855992, linetype=2, colour = "blue") +
  geom_vline(xintercept=1, linetype='longdash') +
  theme_minimal_hgrid(10, rel_small = 1) +
  facet_wrap(~outcome, ncol=1)+
  labs(color = "",y = "", x = "OR",
       title="       MVMR results: mediator with childhood BMI and adult BMI, 95% CI")+
  theme(legend.position = "none", plot.title.position = "plot")
